---
output:
  word_document: default
  html_document: default
  pdf_document: default
---
Exploratory analysis for the term deposits Bank marketing campaign 

#install packages:
# install.packages("ggplot2")
# install.packages("dplyr")
# install.packages("reshape")
# install.packages("corrplot")
# install.packages("rcompanion")
# install.packages("car")
# install.packages("lattice")
# install.packages("vcd")
# install.packages("cramer")
```{r}
library(vcd)
library(corrplot)
library(ggplot2)
library(dplyr)
library(reshape)
library(rcompanion)
library(lattice)
library(car)
library(scales)
library(cramer)
```

###############################################################################
#Upload the Dataset:
```{r}
URL<-("https://raw.githubusercontent.com/MaramShriem/-Marketing-Dataset/main/bank-full.csv")
Dataset<-read.csv(file=URL,header=T,sep=";",na.strings = c(" ","NA"))
```

#Checking NAs':
```{r}
anyNA(Dataset)
#no missing values.
```
#Check duplicates:

```{r}
sum(duplicated(Dataset))
#no duplicate values.
```

#Show the first 6 rows:
```{r}
head(Dataset)
```
#What is the data type, number of columns and number of rows?
```{r}
str(Dataset)
```
#I have 10 categorical variables 7 numeric variables.
```{r}
summary(Dataset)
```
#What are the unique instances for the categorical variables?
```{r}
unique(Dataset$job)
unique(Dataset$marital)
unique(Dataset$education)
unique(Dataset$default)
unique(Dataset$housing)
unique(Dataset$loan)
unique(Dataset$contact)
unique(Dataset$month)
unique(Dataset$poutcome)
unique(Dataset$y)
```
#What is the percentage of customers who will subscribe the product?
```{r}
YesCust<-length(nrow(Dataset)[Dataset$y=="yes"])
round(YesCust/nrow(Dataset)*100,2)
```
#Plot the count of the desired column y.
```{r}
ycol<-ggplot(Dataset, aes(y))
ycol + geom_bar(color = "black",fill = "light blue") + theme(text = element_text(size=10))
```
#or the class distribution 
```{r}
table<-data.frame(prop.table(table(Dataset$y)))
table$Freq<-round((table$Freq)*100,2)
names(table)[names(table) == 'Var1'] <- 'Subscriber'

ggplot(data=table, aes(x=Subscriber, y=Freq)) +
  geom_bar(stat="identity", fill=c("orange","light sky blue"))+
  geom_text(aes(label=Freq), vjust=-0.3, size=3.5)+
  theme_minimal()

#That means the data is imbalanced.
```
#################################################################################
          Analyzing the numeric variables ( we have 7 numeric attributes):
#################################################################################
#What are the means and the means distribution for the numerical independent variables on the ys column:
```{r}
NumIndVar<-Dataset[,c(1,6,10,12,13,14,15,17)]
MD<-data.frame(aggregate(.~y,mean,data=NumIndVar))
MD

dfm<- melt(MD[,-c(8)], id.vars= 1)


ggplot(dfm,aes(x = y, y = value),ylab="Mean") + 
  geom_bar(aes(fill = variable),stat = "identity",position = "dodge")+labs(y= "Mean")
#balance, duration and pdays have changed obviously.
```
#Correlation for numeric variables:
```{r}
DS<-Dataset
DS$y<-ifelse(DS$y=="yes",1,0)

vars  <- c("age","balance","day","duration","campaign","pdays","previous","y") 
m<-DS[vars]

corrplot(cor(m),method="number")

#in General the correlation between the variables is week except there is a slight positive relationship between the y variable( dependent var) and the duration. 
#so let us take a deeper look in this relationship between variable y and the duration.
```
#Analyze the distribution of the duration variable  over the y variable
```{r}
#convert the duration seconds to minutes:
Dataset$durationMin<-  round(Dataset$duration/30)

summary(Dataset$durationMin)
#based on the results i decided to create 5 brackets as i considered the duration over 11 min as an outlier

Dataset$duration_Brkts <- cut(Dataset$durationMin,
                              breaks=c(-1,0,1,3,6,11,164))


Dataset$duration_Brkts<-as.character(Dataset$duration_Brkts)
unique(Dataset$duration_Brkts)

Dataset$duration_Brkts<-ifelse(Dataset$duration_Brkts=="(-1,0]","0",Dataset$duration_Brkts)
unique(Dataset$duration_Brkts)

#getwd()
#write.csv(Dataset,file="Dataset.csv")

#bar plot to display the distribution of duration_Brkts over the class variable.

Dataset$y <- factor(Dataset$y)

other_table<-table(Dataset$duration_Brkts,Dataset$y)

barplot(other_table,
        main = "duration_Brkts Distribution",
        xlab = "y?", ylab = "Frequency",
        col = c("rosybrown", "darkseagreen", "cornflowerblue","royalblue","lavender","thistle"),
        legend.text = rownames(other_table),
        beside = TRUE) # Grouped bars
#or
barplot(other_table,
        main = "duration_Brkts Distribution",
        xlab = "y?", ylab = "Frequency",
        col = c("rosybrown", "darkseagreen", "cornflowerblue","royalblue","lavender","thistle"),
        legend.text = rownames(other_table),
        beside = FALSE) # Stacked bars (default)

##results:
#call duration =0 and (0,1])  means they will not subscribe the product
#most clients who decided to purchase the  product had a call duration between 11-164 minutes.
```
#Correlation test between y and duration variable:
```{r}
cor.test(DS$y,DS$duration)
#the P-value is less than 0.05 which means the relationship is statistically significant. 
```
#Lets check the age distribution:
```{r}
ggplot(Dataset, aes(x=age))+
  ggtitle("age Destribution")+ 
  xlab("age")+ 
  ylab("Density")+
  theme_bw()+#to make the background in a white color
  geom_histogram(aes(y=..density..),binwidth=5,color="light blue",fill='light blue')+
  geom_density(linetype="solid",color="black",adjust=1,size=1)

#test age normality:

#PS: Shapiro function is to test normality of the variable
#(if the the distribution is normal the  P-Value should be greater than 0.05) 

set.seed(10)
x<-sample(Dataset$age,5000)
shapiro.test(x)

#age: not normal distribution and the age is between 30 and 45
```
#Lets check the balance distribution:
```{r}
ggplot(Dataset, aes(x=balance) )+
  ggtitle("balance Destribution")+ 
  xlab("balance")+ 
  ylab("Density")+
  xlim(0,1000)+
  theme_bw()+#to make the background in a white color
  geom_histogram(aes(y=..density..),binwidth=5,color="yellow",fill='#A4A4A4')+
  geom_density(linetype="solid",color="black",adjust=1,size=1)

#test age normality:

#PS: Shapiro function is to test normality of the variable
#(if the the distribution is normal the  P-Value should be greater than 0.05) 

set.seed(10)
x<-sample(Dataset$balance,5000)
shapiro.test(x)

#balance: not normal distribution.
```
#Lets check the day distribution:
```{r}
ggplot(Dataset, aes(x=day) )+
  ggtitle("day Destribution")+ 
  xlab("day")+ 
  ylab("Density")+
  theme_bw()+#to make the background in a white color
  geom_histogram(aes(y=..density..),binwidth=5,color="yellow",fill='#A4A4A4')+
  geom_density(linetype="solid",color="black",adjust=1,size=1)

#test age normality:

#PS: Shapiro function is to test normality of the variable
#(if the the distribution is normal the  P-Value should be greater than 0.05) 

set.seed(10)
x<-sample(Dataset$day,5000)
shapiro.test(x)

#day: not normal distribution.
```
#Lets check the duration distribution:(last call duration)
```{r}
ggplot(Dataset, aes(x=duration) )+
  ggtitle("duration Destribution")+ 
  xlab("duration")+ 
  ylab("Density")+
  xlim(0,350)+
  theme_bw()+#to make the background in a white color
  geom_histogram(aes(y=..density..),binwidth=5,color="yellow",fill='#A4A4A4')+
  geom_density(linetype="solid",color="black",adjust=1,size=1)

#test age normality:

#PS: Shapiro function is to test normality of the variable
#(if the the distribution is normal the  P-Value should be greater than 0.05) 

set.seed(15)
x<-sample(Dataset$duration,5000)
shapiro.test(x)

#duration: not normal distribution.
```
#Lets check the campaign distribution:(campaign is the number of call during this campaign)
```{r}
ggplot(Dataset, aes(x=campaign) )+
  ggtitle("campaign Destribution")+ 
  xlab("campaign")+ 
  ylab("Density")+
  xlim(0,20)+
  theme_bw()+#to make the background in a white color
  geom_histogram(aes(y=..density..),binwidth=5,color="yellow",fill='#A4A4A4')+
  geom_density(linetype="solid",color="black",adjust=1,size=1)

#test age normality:

#PS: Shapiro function is to test normality of the variable
#(if the the distribution is normal the  P-Value should be greater than 0.05) 

set.seed(20)
x<-sample(Dataset$campaign,5000)
shapiro.test(x)

#campaign: not normal distribution.
```
#Lets check the pdays distribution:(pdays is the number of days that passed by after the client was last contacted from a previous campaign)
```{r}
ggplot(Dataset, aes(x=pdays) )+
  ggtitle("pdays Destribution")+ 
  xlab("pdays")+ 
  ylab("Density")+
  xlim(0,75)+
  theme_bw()+#to make the background in a white color
  geom_histogram(aes(y=..density..),binwidth=5,color="yellow",fill='#A4A4A4')+
  geom_density(linetype="solid",color="black",adjust=1,size=1)

#test age normality:

#PS: Shapiro function is to test normality of the variable
#(if the the distribution is normal the  P-Value should be greater than 0.05) 

set.seed(20)
x<-sample(Dataset$pdays,5000)
shapiro.test(x)

#pdays: not normal distribution.
```
#Boxplot to check outliars:
```{r}
boxplot(Dataset$age,outcol="red",pch=23,whiskcol="rosybrown",col="rosybrown",names = "age")
boxplot(Dataset$balance,outcol="red",pch=23,whiskcol="lightblue",col="lightblue",names = "balance",ylim= c(0, 10000))
boxplot(Dataset$day,outcol="red",pch=23,whiskcol="cyan",col="cyan",names = "day")
boxplot(Dataset$duration,outcol="red",pch=23,whiskcol="palegreen",col="palegreen",names = "duration")
boxplot(Dataset$campaign,outcol="red",pch=23,whiskcol="chocolate",col="chocolate",names = "campaign")
boxplot(Dataset$pdays,outcol="red",pch=23,whiskcol="seagreen",col="seagreen",names = "pdays")
boxplot(Dataset$previous,outcol="red",pch=23,whiskcol="seagreen",col="seagreen",names = "previous")
```
         #Categorical variable correlation Matrix#
#Cramer V is Used to calculate the correlation/association between nominal categorical variables.
```{r}
# 0: The variables are not associated
#- 1: The variables are perfectly associated
#- 0.25: The variables are weakly associated
#- .75: The variables are moderately associated

vars  <- c("job","marital","education","default","housing","loan","contact","month","poutcome","y")
df <- Dataset[vars]
# Initialize empty matrix to store coefficients
empty_m <- matrix(ncol = length(df),
                  nrow = length(df),
                  dimnames = list(names(df), 
                                  names(df)))


# Function that accepts matrix for coefficients and data and returns a correlation matrix
calculate_cramer <- function(m, df) {
  for (r in seq(nrow(m))){
    for (c in seq(ncol(m))){
      m[[r, c]] <- assocstats(table(df[[r]], df[[c]]))$cramer
    }
  }
  return(m)
}

cor_matrix <- calculate_cramer(empty_m ,df)

corrplot(cor_matrix,method="number")
#only poutcome and month have a weak association with the dependent variable y.
#month and housing are a very correlated variables.
#education and job are a very correlated variables.
#contact and month are a very correlated variables.
```
#Merge the housing and the loan variables together and recheck the correlation:
```{r}
DS<-Dataset
DS$Totalloans<-ifelse(DS$housing=="yes","yes",ifelse(DS$loan=="yes","yes","no"))
DS$y<-factor(DS$y)
cramerV(DS$Totalloans,DS$y)

Df<-DS
Df$Totalloans<-ifelse(Df$Totalloans=="yes",1,0)
Df$y<-ifelse(Df$y=="yes",1,0)
cor.test(Df$Totalloans,Df$y)

#correlation is negative weak relationship but statistically significant (the Pvalue is less than 0.05)
```
#Marital Status percentages over the desired variable y###
```{r}
DS<-Dataset
DS$y<-ifelse(DS$y=="yes",1,0)

subDiv<-round((aggregate(DS$y,by=list(DS$marital),sum)[1,2]/length(which(DS$y==1))),3)
subMarr<-round((aggregate(DS$y,by=list(DS$marital),sum)[2,2]/length(which(DS$y==1))),3)
subsin<-round((aggregate(DS$y,by=list(DS$marital),sum)[3,2]/length(which(DS$y==1))),3)
subDiv<-label_percent()(subDiv)
subMarr<-label_percent()(subMarr)
subsin<-label_percent()(subsin)

NsubDiv<-round((length(which(DS$marital=="divorced"))-aggregate(DS$y,by=list(DS$marital),sum)[1,2])/length(which(DS$y==0)),4)
NsubMarr<-round((length(which(DS$marital=="married"))-aggregate(DS$y,by=list(DS$marital),sum)[2,2])/length(which(DS$y==0)),4)
Nsubsin<-round((length(which(DS$marital=="single"))-aggregate(DS$y,by=list(DS$marital),sum)[3,2])/length(which(DS$y==0)),4)
NsubDiv<-label_percent()(NsubDiv)
NsubMarr<-label_percent()(NsubMarr)
Nsubsin<-label_percent()(Nsubsin)

ggplot(Dataset,aes(x=y,fill=marital))+
  geom_bar(colour="black",width = .9)+
  theme_classic(base_size = 11)+# background theme
  labs(y="Number of customers",x="y?",title ="marital Status Over ys" )+
  annotate(geom="text",x=2, y=1000, label=subsin)+
  annotate(geom="text",x=2, y=3000, label=subMarr)+
  annotate(geom="text",x=2, y=6000, label=subDiv)+
  annotate(geom="text",x=1, y=5000, label=Nsubsin)+
  annotate(geom="text",x=1, y=20000, label=NsubMarr)+
  annotate(geom="text",x=1, y=37000, label=NsubDiv)
```
#Distribution of the marital status:
```{r}
m<-as.data.frame(table(Dataset$marital))
ycol<-ggplot(m, aes(x=Var1,y=Freq,fill=Var1))
ycol + geom_bar(color = "black",fill = "light gray",stat="identity") +geom_text(aes(label=Freq),vjust=0)
```
#Categorical distribution over the class y
```{r}
attach(Dataset)
ggplot(Dataset, aes(education, ..count..)) + geom_bar(aes(fill = y), position = "dodge")
ggplot(Dataset, aes(marital, ..count..)) + geom_bar(aes(fill = y), position = "dodge")
ggplot(Dataset, aes(job, ..count..)) + geom_bar(aes(fill = y), position = "dodge")
ggplot(Dataset, aes(contact, ..count..)) + geom_bar(aes(fill = y), position = "dodge")
ggplot(Dataset, aes(month, ..count..)) + geom_bar(aes(fill = y), position = "dodge")
ggplot(Dataset, aes(poutcome, ..count..)) + geom_bar(aes(fill = y), position = "dodge")
```
#EDA between some independent variables:
```{r}
ggplot(Dataset, aes(x=marital, y=age)) + geom_boxplot(fill=c('light blue','darkslategray1','turquoise'))
#The plot shows that the average age of unmarried clients is significantly lower than that of the other clients.

ggplot(Dataset, aes(x=age, y=job),width =0.4) + geom_boxplot(fill=c("ivory","cadetblue","honeydew","darkslategray1","darkslateblue","darksalmon","lightcoral","lightcyan3","khaki","lavender","thistle2","turquoise"))
#the age of the most retired customers are between 60 and 27 for students 

ggplot(Dataset, aes(x=y, y=age)) + geom_boxplot(fill=c('light blue','turquoise'))
```



